---
title: "Seurat"
author: "Parissa Amin"
date: "2022-10-04"
output: pdf_document
---



# Creating Seurat object   

Loading packages and data to create Seurat object.

```{r}
library(Seurat) 
library(Matrix)
library(ggplot2)
library(dplyr)
library (patchwork)

metadata <- read.csv("D:/scRNASeq/Data/metadata.csv", header = TRUE, row.names = 1)

gene_names <- read.table("D:/scRNASeq/Data/genes.tsv", header = FALSE, col.names = 
                           "GeneName", stringsAsFactors = FALSE)$GeneName

mtx_matrix<-readMM("D:/scRNASeq/Data/matrix.mtx") 
 
rownames(mtx_matrix) <- gene_names
colnames(mtx_matrix) <- rownames(metadata) 
seurat_obj <- CreateSeuratObject(counts = Matrix::Matrix(as.matrix(mtx_matrix),sparse = T), meta.data = metadata)
```


#Quality Control

While the data is already processed, we perform a brief quality control.

```{r}
# Violin Plot
VlnPlot(object = seurat_obj, features = c('nFeature_RNA','nCount_RNA','pct.Ribo',
                                         'pct.mito'), group.by = 'orig.ident',pt.size = 0,ncol=1)


# FeatureScatter can be used to visualize feature-feature relationships such as number 
#of genes ("nFeature_RNA") vs number of UMIs ("nCount_RNA") 
FeatureScatter(seurat_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA",group.by = 'orig.ident')

```


#Identifying Variable Genes



```{r}
seurat_obj <- FindVariableFeatures(seurat_obj, 
                                   selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(seurat_obj), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seurat_obj)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)

plot1 

plot2
```


#Scaling data

My personal PC doesnt run the full scaling  using 

all.genes  <-rownames(seurat_obj)
seurat_obj <- ScaleData(seurat_obj, features = all.genes)

So I went on with the recommended substitute:

```{r}
seurat_obj <- ScaleData(seurat_obj)
```

#Perform linear dimensional reduction

Next we perform PCA on the scaled data. By default, only the previously determined variable features are used as input, but can be defined using features argument if one wishes to choose a different subset.

```{r}
seurat_obj<-RunPCA(seurat_obj,features =VariableFeatures(object = seurat_obj))
print(seurat_obj[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(seurat_obj, dims = 1:3, reduction = "pca", ncol=3)
DimPlot(seurat_obj, reduction = "pca")
DimHeatmap(seurat_obj, dims = 1, cells = 50000, balanced = TRUE)
ElbowPlot(seurat_obj)
```



#Find clusters

I have tried different values for dim in seurat_obj <-FindNeighbors(seurat_obj, dims = 1:30) and it significantly alters the results! 
I continued using dim=1:30

```{r}
seurat_obj <- FindNeighbors(seurat_obj, dims = 1:30)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.5)
head(Idents(seurat_obj), 5)
```



#UMAP

```{r}
seurat_obj <- RunUMAP(seurat_obj, dims = 1:30)

DimPlot(seurat_obj,reduction="umap")
```

I continued using dim=1:30 here too. 



#Markers for each cluster

Find markers for every cluster compared to all remaining cells, report only the positive ones
 
```{r}
seurat_obj.markers <- FindAllMarkers(seurat_obj, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
seurat_obj.markers %>%
  group_by(cluster) %>%
  slice_max(n = 2, order_by = avg_log2FC)
``` 


Saving the markers: 

```{r}
saveRDS(seurat_obj.markers, file = "seurat_obj.markers.rds")
```

# Conlusion
I was able to identify cellular clusters in scRNA seq atlas of breast tumors. For each cluster, the corresponding marker genes were also identified. The question of whether these findings are biologically valid or not remains to be explored. For example, one needs to explore if the clusters identified here are representing a distinct cell type with their corresponding marker genes.

Additionally, I made multiple decisions about values of parameters that have significantly affected the results. As such, the identified clusters might not represent the real structure of the data and existing clusters (cell types) within it.  

